\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-08-14 papers/methods/paper.tex
Adding challenges section

In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015} to include more robust treatment of systematics and selection effects, whilst improving computational performance. Section \ref{sec:review} is dedicated to a quick review of the supernovae cosmology. In Section \ref{sec:method} we outline our methodology. Model verification on simulated datasets is given in Section \ref{sec:verification}, along with details on potential areas of improvement.

@@

Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements directly. Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. This model is trained separately before fitting the supernovae light curves for cosmology \citep{Guy2010, Mosher2014}. The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics. As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $m_B$, $x_1$ and $c$, where the width of this distribution is called the intrinsic dispersion of the supernova population.

@@

@@

Our second way of addressing the issues with under-reported uncertainty and biased summary statistics are to use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift. {\green This redshift-dependent array is added to our data-set and is included in our modelling by allowing a controlled subtraction of this shift from our observed data, where the amount of shift subtracted is dictated by parameter $s_c$. Specifically, allow the observed colour to shift by $c_{\rm new} = c_{\rm biased} - s_c \times {\rm bias}(z)$. Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a global shift in supernovae colour across all surveys. We place a uniform prior on $s_c$.} With the extra dispersion introduced from the $\kappa$ parameters addressing the under-reporting of error, shifting the mean colour address the non-Gaussianity introduced by chromatic smearing. We allow the amount of shift to be parametrised to account for different intrinsic scatter models where parameter $s_c$ controls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations. The scatter models, sourced from \citet{Guy2010} and \citet{Chotard2011} are discussed in detail in Section \ref{sec:simdes}. The maximum shift was found to using the intrinsic scatter model from \citet{Chotard2011}. It is important to note that the effects being modelled and characterised by simulations here are features of the supernova population, not cosmological features, and so the choice of fiducial cosmology has negligible impact.

@@

With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples that will suffer from significantly more contamination, it will be required that extra degrees of freedom are afforded the outlier population. This contamination is not just a result of non-SNIa being observed, but can also arise from host galaxy misidentification and incorrect redshifts - depending on the cuts on the optimised ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification  \citep{Gupta2016}. Proof of concept analysis on simple simulation data show that an acceptable parametrisation to deal with contamination, both from non type Ia supernovae and from catastrophic failure, is to represent the typically dimmer contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations. For the purposes of the DES spectroscopic sample, which will be confirmed SNIa, $\delta_{M_B}^{\rm outl} =0$. We assume that supernovae fall into the SNIa population as determined by their observed classification probability $\hat{p}$, which is set to unity for the spectroscopic sample.

@@

Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population, but recent analyses require a more sophisticated approach. In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models. The first is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation. The second, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.

>>>>

In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015} to include more robust treatment of systematics and selection effects, whilst improving computational performance. Section \ref{sec:review} is dedicated to a quick review of the supernovae cosmology, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods. In Section \ref{sec:method} we outline our methodology. Model verification on simulated datasets is given in Section \ref{sec:verification}, along with details on potential areas of improvement.

@@

Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements directly. Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. This model is trained separately before fitting the supernovae light curves for cosmology \citep{Guy2010, Mosher2014}. The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$. As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $m_B$, $x_1$ and $c$, where the width of this distribution is called the intrinsic dispersion of the supernova population.

@@

\section{Challenges in Supernova Cosmology}
\label{sec:challenges}

{\red 
The diverse amount of approaches and implementations applied to supernova cosmology is a direct result of the large number of challenges and complications faced by when performing supernova cosmology analyses. In this Section, we outline several of the most prominent challenges.

Forefront among these challenges is our ignorance of the underlying Type Ia population dispersion. Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the variability of Type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements. We utilise two dispersion models in this work. The first is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation. The second, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation. Each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics which is not encapsulated in the reported covariance $\mathcal{C}$. Adopting either scatter model results in markedly different biases, however our uncertainty on the actual dispersion model means we have to support both scatter models.

The underlying supernova population is further complicated by the presence of outliers. Non-Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for Type Ia supernovae. This contamination is not just a result of non-SNIa being observed, but can also arise from host galaxy misidentification causing incorrect redshifts - depending on the cuts on the optimised ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification \citep{Gupta2016} and results in a broad population of outliers. For spectroscopic surveys, where both supernova type and redshift can be confirmed through the supernovae spectra, this outlier population is negligible. However, for photometric surveys which do not have the spectroscopic confirmation, it is one of the largest challenges; how to model, fit and correct for contaminants.

Finally, one of the other persistent challenges facing supernova cosmology analyses are the high number of systematics. Because of the rarity of SNIa, datasets are commonly formed from the discoveries of multiple surveys. However, each different survey introduces additional sources of systematic error, from sources within each survey such as band calibration, to systematics introduced by calibration across surveys. Peculiar velocities, different host environments, dust extinction represent additional sources of systematic uncertainty which must all be modelled and accounted for.
}

@@

Our second way of addressing the issues with under-reported uncertainty and biased summary statistics are to use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift. {\green This redshift-dependent array is added to our data-set and is included in our modelling by allowing a controlled subtraction of this shift from our observed data, where the amount of shift subtracted is dictated by parameter $s_c$. Specifically, allow the observed colour to shift by $c_{\rm new} = c_{\rm biased} - s_c \times {\rm bias}(z)$. Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a global shift in supernovae colour across all surveys. We place a uniform prior on $s_c$.} With the extra dispersion introduced from the $\kappa$ parameters addressing the under-reporting of error, shifting the mean colour address the non-Gaussianity introduced by chromatic smearing. We allow the amount of shift to be parametrised to account for different intrinsic scatter models where parameter $s_c$ controls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations. The maximum shift was found to using the intrinsic scatter model from \citet{Chotard2011}. It is important to note that the effects being modelled and characterised by simulations here are features of the supernova population, not cosmological features, and so the choice of fiducial cosmology has negligible impact.

@@

With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples that will suffer from significantly more contamination, it will be required that extra degrees of freedom are afforded the outlier population. Proof of concept analysis on simple simulation data show that an acceptable parametrisation to deal with contamination, both from non type Ia supernovae and from catastrophic failure, is to represent the typically dimmer contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations. For the purposes of the DES spectroscopic sample, which will be confirmed SNIa, $\delta_{M_B}^{\rm outl} =0$. We assume that supernovae fall into the SNIa population as determined by their observed classification probability $\hat{p}$, which is set to unity for the spectroscopic sample.

@@

Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population, but recent analyses require a more sophisticated approach. In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.
